Synthesizing Symmetric Lenses

### 2018-07-10 papers/SymmetricLenses/main.tex

(The idea of lenses.  Their usefulness and popularity.)

Recent work by~\citet{optician} has also demonstrated
how techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize the simplest class of lenses: so-called {\em
  bijective lenses}, where no information is discarded when transforming
data in either direction.

We show here how to extend this algorithm to a much richer class of lenses,
which we call {\em simple symmetric lenses}; this class includes all
bijective and standard ``asymmetric'' lenses~\cite{OriginalFosterPOPLpaper},
as well as a rich subset of the full-blown ``symmetric lenses'' studied by
\citet{symmetric-lenses}.

(why our algorithm is harder; how we use information theory to help)

\end{abstract}

>>>>

A {\em lens} is a program that can be run both ``front to back'' and ``back
to front,'' allowing data to be transformed in two directions.  Since their
introduction by~\citet{Focal2005-long2}, lenses have been extensively
studied and applied.  Recent work~\cite{optician} has also demonstrated how
techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize a very simple class of lenses---so-called {\em
  bijective lenses} over string data---given a pair of types (regular
expressions) and a small number of examples.

We show how to extend this synthesis algorithm to a much richer class of
lenses, which we call {\em simple symmetric lenses}, including both
bijective lenses and the more widely used ``asymmetric''
lenses~\cite{Focal2005-long2}, as well as a rich subset of the full-blown
``symmetric lenses'' proposed by \citet{symmetric-lenses}.  Simple symmetric
lenses are of independent theoretical interest, being the largest class of
symmetric lenses that do not rely on persistent internal state.

Synthesizing simple symmetric lenses is substantially more challenging that
synthesizing bijective lenses: since some of the information on each side
can be ``disconnected'' from the other side, there will in general be {\em
  many} lenses that agree with a given example.  To guide the search
process, we use {\em stochastic regular expressions} and ideas from
information theory to estimate the amount of information propagated by a
candidate lens.

We describe an implementation of simple symmetric lenses and our synthesis
procedure as extensions to the Boomerang language~\cite{Boomerang} and
evaluate its performance on a range of benchmark examples.
\end{abstract}
